### 2021-07-29 presentations/FOSS4G2021_presentation.Rmd
FOSS4G presentation structure

---
title: "ONE ARABESQUE IN THE SMALL WORLD OF OD WEBMAPS"
subtitle: "FOSS4G 2021"  
author: 
  - "Nicolas Roelandt"
date: '`r Sys.Date()`'
output:
  xaringan::moon_reader:
    seal: false
    css: xaringan-themer.css
    nature:
      slideNumberFormat: "%current%"
      highlightStyle: github
      highlightLines: true
      ratio: 16:9
      countIncrementalSlides: true
---
```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(
  fig.width=9, fig.height=3.5, fig.retina=3,
  out.width = "100%",
  cache = FALSE,
  echo = TRUE,
  message = FALSE, 
  warning = FALSE,
  hiline = TRUE
)
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_duo_accent(
  primary_color = "#2f2a85",
  secondary_color = "#2f2a85",
  inverse_header_color = "#ef7d00"
)
```
class: inverse center middle

<style>
.top_left {
    left: 0;
    position: absolute;
    top: 0;
}
.top_left img {
margin: 0;
}

.top_right {
    right: 0;
    position: absolute;
    top: 0;
}
.top_right img {
margin: 0;
}
</style>

.top_left[
```{r logo, echo=FALSE, out.width="100%", eval=TRUE}
knitr::include_graphics("media/logo_FOSS4G2021.svg")
```
]

.top_right[
```{r logo2, echo=FALSE, out.width="100%", eval=TRUE}
knitr::include_graphics("media/logo_univ_gustave_eiffel_blanc_rvb.svg")
```
]

# ONE ARABESQUE IN THE SMALL WORLD OF OD WEBMAPS

## <span style='color:#ef7d00;'>_FOSS4G 2021 - remote_</span>

### Nicolas Roelandt
---

# Introduction

## Contexte

--
- Focus on web applications

--
- Thematic geovisualization and cartographic purposes

--
- Flows and spatial networks

--
- Review of existing tools (2016 -)

--
- Survey within 200 users

--
- Development of Arabesque (2018 -)

???

The paper focuses on a specific part of geoweb, it relates to flows and spatial networks for thematic geovisualization and cartographic purposes.

Arabesque, after an online survey of people interested in geovisualizing this type of data and a review of existing web pages and applications.

Particularities of flow mapping : pay close attention to the preparation and processing of the OD data before representation, otherwise graphical (e.g. spaghetti effect) and visual complexities will arise.

arabesque positioning: combining OD data visualization and cartography (not also mapping).

3 key points: dealing with numerical data, geographical data, display.

---
class: inverse center middle

# Why compare ?
---
layout: true
# Why compare ?
---

- Arabesque V1 available, V2 in dev

--
- Is it really innovative ?

--
- Help users to make an informed choice

???

Arabesque is now an usable application, useable by everyone.
It is not perfect, some features we think important are missing, there is inconvenient bugs.
So we are working on a version 2.

But is Arabesque really innovative, does it brings something to the community that are not in similare tools ?

So we made this comparison and wrote this paper to help user make an informed choice

---
layout: false
class: inverse center middle
# Comparison criteria

---
layout: true
# Comparison criteria

---

toto
---
layout: false
class: center, middle

# Thanks!

Nicolas Roelandt - Univ. Gustave Eiffel

[nicolas.roelandt@univ-eiffel.fr](mailto:nicolas.roelandt@univ-eiffel.fr) [@RoelandtN42](https://twitter.com/RoelandtN42)

.left[

##### Slides created via the R packages:

- [**xaringan**](https://github.com/yihui/xaringan)<br>
- [gadenbuie/xaringanthemer](https://github.com/gadenbuie/xaringanthemer)
]