### 2021-07-29 presentations/FOSS4G2021_presentation.Rmd
Complete structure

---
title: "ONE ARABESQUE IN THE SMALL WORLD OF OD WEBMAPS"
subtitle: "FOSS4G 2021"  
author: 
  - "Nicolas Roelandt"
date: '`r Sys.Date()`'
output:
  xaringan::moon_reader:
    seal: false
    css: ["xaringan-themer.css", "custom.css"]
    nature:
      slideNumberFormat: "%current%"
      highlightStyle: github
      highlightLines: true
      ratio: 16:9
      countIncrementalSlides: true
---
```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(
  fig.width=9, fig.height=3.5, fig.retina=3,
  out.width = "100%",
  cache = FALSE,
  echo = TRUE,
  message = FALSE, 
  warning = FALSE,
  hiline = TRUE
)
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_duo_accent(
  primary_color = "#2f2a85",
  secondary_color = "#2f2a85",
  inverse_header_color = "#ef7d00"
)
```

class: inverse center middle title-slide

.top_left[
```{r logo, echo=FALSE, out.width="100%", eval=TRUE}
knitr::include_graphics("media/logo_FOSS4G2021.svg")
```
]

.top_right[
```{r logo2, echo=FALSE, out.width="100%", eval=TRUE}
knitr::include_graphics("media/logo_univ_gustave_eiffel_blanc_rvb.svg")
```
]

# ONE ARABESQUE IN THE SMALL WORLD OF OD WEBMAPS

## <span style='color:#ef7d00;'>_FOSS4G 2021 - remote_</span>

### Nicolas Roelandt
---

# Introduction

## Contexte

--
- Focus on web applications

--
- Thematic geovisualization and cartographic purposes

--
- Flows and spatial networks

--
- Review of existing tools (2016 -)

--
- Survey within 200 users

--
- Development of Arabesque (2018 -)

???

The paper focuses on a specific part of geoweb, it relates to flows and spatial networks for thematic geovisualization and cartographic purposes.

Arabesque, after an online survey of people interested in geovisualizing this type of data and a review of existing web pages and applications.

Particularities of flow mapping : pay close attention to the preparation and processing of the OD data before representation, otherwise graphical (e.g. spaghetti effect) and visual complexities will arise.

arabesque positioning: combining OD data visualization and cartography (not also mapping).

3 key points: dealing with numerical data, geographical data, display.

---
class: inverse center middle

# Why compare ?
---
layout: true
# Why compare ?
---

- Arabesque V1 available, V2 in dev

--
- Is it really innovative ?

--
- Help users to make an informed choice

???

Arabesque is now an usable application, useable by everyone.
It is not perfect, some features we think important are missing, there is inconvenient bugs.
So we are working on a version 2.

But is Arabesque really innovative, does it brings something to the community that are not in similare tools ?

So we made this comparison and wrote this paper to help user make an informed choice

---
layout: false
class: inverse center middle
# Comparison criteria

---
layout: true
# Comparison criteria

---
Applications similar to Arabesque :
  * Web applications
  * FOSS
  * OD data visualisation capabilities

???

Opportunities for action on the data and their representation, export, etc. 
Types of cartographic rendering
---
layout: false
class: inverse center middle
# Data model

---
layout: true
# Data model
---

* data format may varie

--
* data stockage may varie : wide/large matrix format

--
* nodes coordinates in WGS84 (lat/lon)

--
* no cleaning tools provided

???
Data format: 2 files nodes / links is the most common (all but Kepler.gl).
Cleaning:
Your data needs to be prepared ahead using other tools.
Arabesque provide information on rejected data.

---
layout: false
class: inverse center middle
# Usage

---
layout: true
# Usage
---
* Data Analysis
  * Arabesque only

--
* Exploration
  * Arabesque
  * KeplerGL
  * Flowmap.blue

???
Only Arabesque offers real data analysis capabilities with custom filtering options and new indicators computation.

---

* Representation
  * Arabesque
  * KeplerGL
  * Flowmap.blue
  * Magrit

???
All applications can be used to represent OD data.
However Kepler.gl and Flowmap.blue offers limited design options (3D arcs).
Magrit is dedicated to thematic mapping

---
layout: false
class: inverse center middle
# Public/Purpose

---
layout: true
# Public/Purpose
---

* Data exploration and visualisation
    * data journalist
    * general public
    * web publishing

* Academia and teaching
    * thematic maps production
    * statistical analysis capabilities
    * Arabesque and Magrit

---
layout: false
class: inverse center middle
# Performance

---
layout: true
# Performance
---

* Technology dependant
  * KeplerGl and Flowmap.blue
    * Deck.gl
    * large datasets
    * fast
    * limited graphic options
  * Arabesque and Magrit
    * D3
    * smaller datasets (< 10 000 objects)
    * more graphic options

---
layout: false
class: inverse center middle
# Interoperability and reproducibility

---
layout: true
# Interoperability and reproducibility
---

## data format

  * CSV and/or GeoJSON
  * 2 separates tables (except for KeplerGL)
  * for Arabesque, Magrit and Flowmap.blue, you can use the same files
      * note that for Flowmap.blue you will have to save the CSV file in a google sheet document or copy-paste the data in a browser

---
## Export

  • online hosting (KeplerGL, Flowmap.blue)
  • image export (Arabesque, Magrit)
  • Data and project parameters (Arabesque, KeplerGl, Magrit)

---
layout: false
class: inverse center middle
# Conclusion

---
layout: true
# Conclusion
---  
# Conclusion

- business culture, user's skills and final use of the map influence the choice of the tool because it will influence the chosen approaches (exploration and/or analysis)
- data does not influence the choice of the tool
- whole range of needs and possibilities are covered, for the benefits of the community.
layout: false
class: center, middle

# Thanks!

Nicolas Roelandt - Univ. Gustave Eiffel

[nicolas.roelandt@univ-eiffel.fr](mailto:nicolas.roelandt@univ-eiffel.fr) [@RoelandtN42](https://twitter.com/RoelandtN42)

.left[

##### Slides created via the R packages:

- [**xaringan**](https://github.com/yihui/xaringan)<br>
- [gadenbuie/xaringanthemer](https://github.com/gadenbuie/xaringanthemer)
]